---
title: "Test Geocoders"
author: "Branson Fox"
date: '(`r format(Sys.time(), "%B %d, %Y")`)'
output: 
  github_document: default
  html_notebook: default
---

## Introduction
This notebook uses 7 different available services to geocode data, measuring both the response and the accuracy of the response.

## Dependencies
These packages are required for the analysis in this notebook.
```{r dependencies}
library(httr)
library(dplyr)
library(glue)
library(magrittr)
library(Imap)
library(yaml)
library(censusxy)
```

### Load Geocoders
We will also load the geocoding implementations we wrote.
```{r load-geocoders}
# Single Line Geocoders
source('../source/singleLine.R')
# Batch Geocoders
source('../source/bing.R')
source('../source/esri.R')
source('../source/geocodio.R')
source('../source/here.R')
source('../source/tomtom.R')
```

### Load Credentials
We need to load the API-Keys for these services
```{r keys}
creds = yaml::read_yaml('../creds.yml')
```

## Load Data
Next, we'll load the data to be geocoded. We need to add the city/state suffix.
```{r load-data}
# Load the Ground Truth
local <- read.csv('../data/STL_CRIME_Homicides_local_id.csv', stringsAsFactors = FALSE)
addresses <- read.csv('../data/STL_CRIME_Homicides.csv', stringsAsFactors = FALSE)$address_norm %>%
  paste0(' St. Louis, MO')
```

## Execute Geocoders
Now, we'll run the geocoders and record a time.
```{r geocode}
# Single Line ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
Esri_Single_Time <- system.time({
  Esri_Single <- batch(addresses, ArcMap, token = creds$esri)
})
save(Esri_Single, Esri_Single_Time, file = '../results/Esri_Single.rda')
#
Google_Single_Time <- system.time({
  Google_Single <- batch(addresses, GoogleMaps, key = creds$google)
})
save(Google_Single, Google_Single_Time, file = '../results/Google_Single.rda')
#
HERE_Single_Time <- system.time({
  HERE_Single <- batch(addresses, HERE, key = creds$here)
})
save(HERE_Single, HERE_Single_Time, file = '../results/HERE_Single.rda')
#
OpenCage_Single_Time <- system.time({
  OpenCage_Single <- batch(addresses, OpenCage, key = creds$opencage)
})
save(OpenCage_Single, OpenCage_Single_Time, file = '../results/OpenCage_Single.rda')
#
TomTom_Single_Time <- system.time({
  TomTom_Single <- batch(addresses, TomTom, key = creds$tomtom)
})
save(TomTom_Single, TomTom_Single_Time, file = '../results/TomTom_Single.rda')
#
Geocodio_Single_Time <- system.time({
  Geocodio_Single <- batch(addresses, Geocodio, key = creds$geocodio)
})
save(Geocodio_Single, Geocodio_Single_Time, file = '../results/Geocodio_Single.rda')
#
Census_Single_Time <- system.time({
  Census_Single <- batch(addresses, CensusBureau)
})
save(Census_Single, Census_Single_Time, file = '../results/Census_Single.rda')
#
Bing_Single_Time <- system.time({
  Bing_Single <- batch(addresses, Bing, key = creds$bing)
})
save(Bing_Single, Bing_Single_Time, file = '../results/Bing_Single.rda')

# Batch ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
Bing_Batch_Time <- system.time({
  Bing_Batch <- bing(addresses, creds$bing)
})
save(Bing_Batch, Bing_Batch_Time, file = '../results/Bing_Batch.rda')
#
Esri_Batch_Time <- system.time({
  Esri_Batch <- esri(addresses, creds$esri)
})
save(Esri_Batch, Esri_Batch_Time, file = '../results/Esri_Batch.rda')
#
TomTom_Batch_Time <- system.time({
  TomTom_Batch <- tomtom(addresses, creds$tomtom)
})
save(TomTom_Batch, TomTom_Batch_Time, file = '../results/TomTom_Batch.rda')
#
Geocodio_Batch_Time <- system.time({
  Geocodio_Batch <- geocodio(addresses, creds$geocodio)
})
save(Geocodio_Batch, Geocodio_Batch_Time, file = '../results/Geocodio_Batch.rda')
#
Census_Batch_Time <- system.time({
  # get usable data format
  df <- read.csv('../data/STL_CRIME_Homicides.csv', stringsAsFactors = FALSE) %>%
    mutate(state = 'MO', city = 'St. Louis')
  Census_Batch <- censusxy::cxy_geocode(df, address_norm, city, state, output = 'tibble')
})
save(Census_Batch, Census_Batch_Time, file = '../results/Census_Batch.rda')
#
HERE_Batch_Time <- system.time({
  HERE_Batch <- here(addresses, creds$here)
})
save(HERE_Batch, HERE_Batch_Time, file = '../results/HERE_Batch.rda')
```

### Build a Dataframe
We'll now build a large dataframe with all of our results
```{r build dataframe}
single_to_col <- function(list, prefix){
  lats = sapply(list, function(x){tryCatch(x$lat, error = function(e) return(NA) )})
  lons = sapply(list, function(x){tryCatch(x$lon, error = function(e) return(NA) )})
  df = data.frame(stringsAsFactors = FALSE,
                  lat = lats,
                  lon = lons)
  
  names(df) <- paste0(prefix, '.', names(df))
  return(df)
}

all_geocodes <- 
  data.frame(stringsAsFactors = FALSE,
    address = paste0(local$address_norm, ' St. Louis, MO'),
    truth.lat = local$gw_y,
    truth.lon = local$gw_x,
    single_to_col(Bing_Single, 'BingSingle'),
    single_to_col(Esri_Single, 'EsriSingle'),
    single_to_col(Google_Single, 'Google'),
    single_to_col(HERE_Single, 'HereSingle'),
    single_to_col(OpenCage_Single, 'OpenCage'),
    single_to_col(TomTom_Single, 'TomTomSingle'),
    single_to_col(Census_Single, 'CensusSingle'),
    single_to_col(Geocodio_Single, 'GeocodioSingle')
)

# TomTom
all_geocodes %<>% 
  left_join(
    TomTom_Batch %>% 
      filter(!duplicated(TomTom_Batch$address)) %>%
      rename(TomTomBatch.lat = lat,
             TomTomBatch.lon = lon)
    )
# Bing
all_geocodes %<>%
  left_join(
    Bing_Batch %>%
      filter(!duplicated(Bing_Batch$GeocodeRequest.Address.AddressLine)) %>%
      transmute(address = GeocodeRequest.Address.AddressLine %>% 
                  gsub('St. Louis','St. Louis, MO',.),
                BingBatch.lat = GeocodeResponse.Point.Latitude,
                BingBatch.lon = GeocodeResponse.Point.Longitude)
  )

# (Ordered Batches)
all_geocodes %<>%
  cbind(
    Census_Batch %>%
      transmute(CensusBatch.lat = lat,
                CensusBatch.lon = lon),
    Geocodio_Batch %>%
      transmute(GeocodioBatch.lat = lat,
                GeocodioBatch.lon = lng),
    Esri_Batch %>%
      transmute(EsriBatch.lat = y,
                EsriBatch.lon = x),
    HERE_Batch %>%
      transmute(HereBatch.lat = displayLatitude,
                HereBatch.lon = displayLongitude)
    )
```

## Compare Timing
To compare the timing of these geocoders:
(These were ran multiple times with little variability between trials)
```{r timing}
# Get and Sort Fastest Times
times <- function(){
  time_objs <- ls(envir = .GlobalEnv) %>%
    grep('Time' ,.,value = TRUE)
  
  time <- mget(time_objs, envir = .GlobalEnv)
  results = vector('numeric', length(time))
  
  for (i in seq_along(results)){
    results[i] <- time[[i]][3]
  }
  names(results) <- names(time)
  
  results %<>% sort
  
  print(results)
}
times()
```

## Compare Match Rate
This is the number of responses out of 1822 addresses submitted. Note, any valid response is counted, regardless of accuracy.
```{r match-rate}
gmr <- function(vector){
  1822 - sum(is.na(vector))
}

lats = names(all_geocodes)[seq(from = 2, to = 30, by = 2)]
for (i in lats){
  print(i %>% gsub('.lat','',.))
  print(gmr(all_geocodes[i]))
}
```
Note that our ground truth is not as thorough as the commerical geocoders even, which is a serious limitation. However, we will filter the data in order to accomplish a fair comparison of known locations.

## Compare Accuracy
In order to compare accuracy, we first calculate the difference in distance from the ground truth of each point (in Meters) and then calculate the root mean square error (RMSE) for all of the points.

First, we'll implement a function to calculate the distance difference, and then a fucntion for calculation root mean square error.

The formulat for RMSE is as follows:
$RMSE = \sqrt{\frac{1}{n}\sum_{t=1}^{n}e_t^2}$

```{r funcs}
dist_diff <- function(expected_lon, expected_lat, actual_lon, actual_lat){
  # Accepts Vector of Expected Lat/Lon
  # And Vector of Actual Lat/Lon
  # Replace NAs
  actual_lon %<>% sapply(function(x){
    if(is.na(x)){
      return(0)
    }else{
      return(x)
    }
  })
  actual_lat %<>% sapply(function(x){
    if(is.na(x)){
      return(0)
    }else{
      return(x)
    }
  })
  
  diffs <- vector('numeric', length(expected_lon))
  for (i in seq_along(diffs)){
    diffs[i] <- Imap::gdist(expected_lon[i], expected_lat[i], actual_lon[i], actual_lat[i], units = 'm')
  }
  # Return Vector of Distance Difference
  return(diffs)
}

rmse <- function(vector){
  # Accepts a Vector of Differences from Expected
  n = length(vector)
  sq = vector ^ 2
  sm = sum(sq/n)
  root = sqrt(sm)
  return(root)
}

```
We also need to create a dataset that we can fairly compared to the ground truth, so we'll remove the NAs from our ground truth.

```{r}
truth <- dplyr::filter(all_geocodes, !is.na(all_geocodes$truth.lat))
```

Now to run the analysis of accuracy:
```{r}
truth.lat = truth$truth.lat
truth.lon = truth$truth.lon
lats = names(truth)[seq(from = 4, to = 30, by = 2)]
lons = names(truth)[seq(from = 5, to = 31, by = 2)]

for (i in seq_along(lats)){
  res = dist_diff(truth.lon, truth.lat, truth[[lons[i]]], truth[[lats[i]]]) %>%
    rmse
  
  print(lats[i] %>% gsub('.lat','',.))
  print(res)
}

# Removing the NAs instead of Making 0.... Otherwise radical census results
```

